---
title: "Multiple Coordinated Views<br>with `plotly` and `crosstalk`"
author: "Chen Dong"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(reshape2)
library(plotly)
library(crosstalk)
```

## Libraries 

The code used to create the plots below requires inclusion of the following libraries:

  * `tidyverse`
  * `reshape2`
  * `plotly`  
  * `crosstalk`

## Session Information

```{r, echo=FALSE}
sessionInfo()
```

```{r, echo=FALSE}
# read and prepare the data
st <- read.table('flunet2010_11countries.csv', sep=',', header=T, stringsAsFactors=F, skip=1)

# additional data to be used for some visualizations
d1 <- data.frame(country=c("Afghanistan","Argentina",
                        "Australia","Canada", "China", 
                        "Colombia", "Egypt", "Germany", 
                        "Ireland", "South.Africa", "USA"),
                 continent=c('Asia', 'South America', 'Australia', 'North America', 
                        'Asia', 'South America', 'Africa', 'Europe', 'Europe', 
                        'Africa', 'North America'),
                 hemisphere=c('Northern', 'Southern', 'Southern', 'Northern', 'Northern', 
                        'Northern', 'Northern', 'Northern', 'Northern', 'Southern', 
                        'Northern'),
                 code=c("AFG","ARG",
                        "AUS","CAN", "CHN", 
                        "COL", "EGY", "DEU", 
                        "IRL", "ZAF", "USA"),
                 stringsAsFactors=FALSE)

# reshape table from wide to long format
lst <- melt(st, id.vars=1:1, variable.name="country", value.name="count")

# map additional data to data frame
lst$continent <- d1$continent[match(lst$country, d1$country)]
lst$hemisphere <- d1$hemisphere[match(lst$country, d1$country)]
lst$code <- d1$code[match(lst$country, d1$country)]
```

## Visualization 1

#### Create Linked Plots Showing Flu Cases per Country and Total Flu Cases per Week

In this example a dropdown selector and a time slider provided by the _crosstalk_ package are linked to two plots created by _plotly_.

Key methods and settings used:

* `SharedData$new` (`crosstalk` package) to create a new shared data object that can represent a collection of selected objects, e.g. to enable brushing and linking across plots
* `add_markers`, `add_lines`, `add_bars` to create the corresponding plots
* `layout( barmode = 'stack')` to create a stacked bar chart
* `subplot` (`plotly` package) to combine two or more `plotly` plots with `shareX=TRUE` to combine the `x` axes of the plots
* `filter_select` and `filter_slider` to create `crosstalk` controls
* `bscols` (`crosstalk` package) to arrange plots and controls into columns
* `hide_legend` to turn off Plotly legend

#### Recommended Resources

* https://plotly-book.cpsievert.me/linking-views-without-shiny.html
* https://plot.ly/r/bar-charts/
* https://rstudio.github.io/crosstalk/using.html

******

```{r, echo=FALSE, warning=FALSE}
sd1 <- SharedData$new(lst)
base1 <- plot_ly(sd1) 

p1 <- 
  base1 %>%
  hide_legend() %>%
  add_lines( x=~week, y=~count, color=~country, colors = "Dark2") %>%
  add_markers( x=~week, y=~count, color=~country, colors = "Dark2")

# 1. create a stacked bar chart
p2 <- 
  base1 %>% 
  hide_legend() %>%
  add_bars(x=~week, y=~count,  colors = "Dark2") 

# 2. create a subplot to link the two plots
p3 <- subplot(p1, p2, titleX = TRUE, nrows = 2, shareX = TRUE) %>% 
  layout( barmode = 'stack', title = "Flu Cases by Country and Totals week") %>%
  hide_legend() 

# 3. add the slider control to filter weeks to this bscols configuration
bscols(
  widths=c(3,9),
  list(
    filter_select('count', 'Country', sd1, ~country ),
    filter_slider('week','Week',sd1,~week)),
  p3) 


# 4. replace this plot with your subplot object
```

******

## Visualization 2

#### Create an Overview+Detail Plot Showing Flu Cases per Country

In this example a checkbox selection is used to control countries of which continents are shown in a stacked bar chart. The stacked bar chart shows flu cases per week and country. The overview+detail visualization is enabled by using the `rangeslider` property on the `x` axis of the plot.

Key methods and settings used:

* `filter_checkbox` to create `crosstalk` checkbox list. 

#### Recommended Resources

* https://plot.ly/r/range-slider/

******

```{r, echo=FALSE, warning=FALSE}
sd2 <- SharedData$new(lst)
base2 <- plot_ly(sd2) 

p4 <- 
  base2 %>% 
  hide_legend() %>%
  add_bars(x=~week, y=~count, color=~country ,colors = "Dark2") %>%
  add_markers(x=~week, y=~count, color=~country ,colors = "Dark2") %>%
  layout(barmode = 'stack', title = "Flu Cases by Country", xaxis = list(rangeslider = list(type = "week")))
  

bscols(
  widths=c(3,9),
  list(
    filter_checkbox('continent', 'Continent', sd2, ~continent )),
  p4) 

```

******

## Visualization 3

#### Create Linked Plots Showing Flu Cases per Country per Week and Total Flu Cases per Country

For this visualization we create two linked plots. One that shows flu cases per country per week and a second on that show the total of all flu cases per country.

Key methods and settings used:

* `group_by(country)` to group the data frame by country
* `summarise` with `sum` to compute the totals per country
* `arrange` to sort the countries by the total of flu cases
* `subplot` to link the two plots

#### Recommended Resources

* https://plotly-book.cpsievert.me/linking-views-without-shiny.html (in particular _4.2.5 The SharedData plot pipeline_)

******

```{r, echo=FALSE, out.width="100%"}
sd4 <- SharedData$new(lst, key=~country)
base4 <- plot_ly(sd4) %>% group_by(country)

p5 <- base4 %>% 
  hide_legend() %>%
  add_lines(x=~week, y=~count, color=~country ,colors = "Dark2") %>%
  layout( title = "Flu Cases by Country and Country Total")


p6 <- base4 %>% 
  hide_legend() %>%
  summarise(count.sum=sum(count,na.rm = T)) %>%
  arrange(desc(count.sum)) %>% 
  add_bars(x=~count.sum, y=~factor(country, levels = country), colors = "Dark2")


s4 <- subplot(p5, p6, titleX = TRUE, nrows = 2, shareX = F) %>% 
  hide_legend() 

s4
```

******

## Visualization 4

#### Create a Choropleth Map Showing Flu Cases per Country

This visualization uses `crosstalk` controls to select continent and weeks for which flu cases should be shown in the choropleth map. 

Key methods and settings used:

* The appearance of the map is controlled by `showframe`, `showcoastlines`, `showland`, `landcolor`, and `projection` (here 'Mercator').

#### Recommended Resources

* https://plot.ly/r/choropleth-maps/

******

```{r, echo=FALSE, warning=FALSE}
sd3 <- lst %>% na.omit() %>% SharedData$new()

g <- list(
  showframe = F,
  showcoastlines = T,
  showland = T,
  landcolor = toRGB("grey"),
  projection = list(type = 'Mercator')
)

p7 <- plot_geo(sd3) %>% 
  add_trace(z=~count,color = ~count, colors = 'Blues', text = ~country, locations = ~code, 
            marker = list(line = list(color = toRGB("grey"), width = 0.5))) %>%
  colorbar(title = "Flu Cases") %>%
  layout(
    title = 'Flu Cases',
    geo = g
  )

bscols(
  widths=c(3,9),
  list(
    filter_checkbox('continent', 'Continent', sd3, ~continent ),
    filter_select('week','Week',sd3,~week,multiple = F)),
  p7) 
```
