---
title: "R Shiny Example"
author: "Chen Dong"
output: html_document
runtime: shiny
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reshape2)
library(plotly)
library(shiny)
library(network)
library(tidyverse)
```

## Libraries 

The code used to create the plots below requires inclusion of the following libraries:

  * `reshape2`
  * `plotly`  
  * `shiny`

## Overview

In this notebook, we will demonstrate how to use an R Shiny server to provide interaction with a plot.

First we have to load our data and reshape it so that the country is in a column called `country`. So we go from a table with countries as the columns:

```
week Afghanistan Germany
1           10      60
2           15      63
3           25      64
```

To one where the country and counts are in their own column:

```
week  country count
1     Afghanistan 10
2     Afghanistan 15
3     Afghanistan 25
1     Germany     60
2     Germany     63
3     Germany     64
```

This will make it easier to filter the data on the Shiny server.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
t <- read.table('../HW1/flunet2010_11countries.csv', sep=',', header=T, stringsAsFactors=F)

head(t)

library(shiny)
library(reshape2)
library(plotly)

d1 <- data.frame(Col1=c("Afghanistan","Argentina",
                        "Australia","Canada", "China", 
                        "Colombia", "Egypt", "Germany", 
                        "Ireland", "South.Africa", "USA"),
                 Col2=c('Asia', 'South America', 'Australia', 'North America', 
                        'Asia', 'South America', 'Africa', 'Europe', 'Europe', 
                        'Africa', 'North America'),
                 stringsAsFactors=FALSE)

colnames(t) <- t[1,]
t <- t[-1,]

lst <- melt(t, id.vars=1:1, 
            variable.name="country", 
            value.name="count")

lst$week <- as.numeric(lst$week)
lst$count <- as.numeric(lst$count)
```

Next we will create an R Shiny app that lets us select the country and weeks that are displayed in our plot. The UI element will contain a selection box for the country and an input slider for the weeks. This control will pass its values to the server which will return a rendered plot.

The interface between the UI and server is through the `input` variable. Each selection control in the UI sets a member of the input variable which can then be accessed in the server. As part of the problem set, we will add a selector that lets us choose which country we wish to view the plot for (Problem 1.1), a slider that lets us select a week range (Problem 1.2) and an event handler that lets us update the UI controls in response to user interactions with the Plotly plot (Problem 1.3).

```{r, echo=FALSE, warning=FALSE}
ui <- fluidPage(
  titlePanel("Flu across the world"),
  sidebarPanel(
    # 1.1 Add a dropdown for the list of countries that we can select
    selectInput("select", 
                label = "Select a country:", 
                choices = d1$Col1,
                selected = "Afghanistan"),
    # 1.2 Add a slider to select which weeks should be visible
    #)
    sliderInput("slide", 
                label = "Range:",
                min = 1, max = 52, value = c(1, 52))
  ),
  
  mainPanel(
    plotlyOutput("plot"),
    
    # we need the following output so that we catch events in
    # the plots
    
    verbatimTextOutput("click")
  )
);

server <- function(input, output,session) {
  output$plot <- renderPlotly({

    output$plot <- renderPlotly({
      # (1.1 and 1.2) Display the line plot as filtered by the input country
      # and week range
      dat <- lst[lst$country==input$select,]
      p <- ggplot(dat) + geom_line(aes(x=week,y=count)) + xlim(input$slide)
      ggplotly(p)
    
    })
    
    output$click <- renderPrint({
      # Problem 1.3, update the week slider here
      d <- event_data("plotly_relayout")
      updateSliderInput(session, "slide", value = round(as.numeric(d))[1:2])
    });
  })
};

shinyApp(ui, server, options = list(height=600));
```

## Example 2

This example will illustrate two linked plots as well as a controller that modifies both. The purpose of these plots is to show the correlation in flu cases between different countries. The first plot is a heatmap of absolute values of the pairwise correlation coefficients. The second is a node-link diagram showing which countries have highly correlated flu incidence rates.

Because the threshold for the correlation that we use to create the node link diagram is somewhat arbitrary, we will include a slider that we can use to vary it and see its effect on the heatmap and node-link (Problem 2.1). To help us get a sense of which cells in the matrix correspond to which links in the diagram, we will add an interaction that highlights links when we click on a cell in the matrix (Problem 2.2). While working on this problem, beware of two caveats:

1. Because the matrix is filtered to contain only the correlations that are above our threshold, clicking on cells with a lower threshold will not highlight anything. 

2. As far as we can tell, plotly does not provide a way to click on lines connecting points so the interaction will be unidirectional.

```{r, echo=FALSE, warning=FALSE}
countries <- unique(lst$country)

# calculate the correlation matrix
# without using the "week" columns
t <- as.data.frame(data.matrix(t))

tc <- t[ , !(names(t) %in% c("week"))]
m <- abs(cor(tc))

ui <- fluidPage(
  titlePanel("Correlations between flu incidences in different countries"),
  
  # Problem 2.1, add a control for selecting the correlation coefficient
  sidebarPanel(
    sliderInput("slide",
                label = "Correlation coefficient threshold:",
                min = 0,
                max = 1,
                value = 0.5)
  ),
  
  mainPanel(
    plotlyOutput("plot"),
    plotlyOutput("network"),
    verbatimTextOutput("click"),
    verbatimTextOutput("event")
    
  )
);

server <- function(input, output,session) {
  output$network <- renderPlotly({
    set.seed(1);
    
    tc <- t[ , !(names(t) %in% c("week"))]
    m <- abs(cor(tc))
    m2 <- m
    m2[abs(m)<input$slide] <- 0
    
    # Problem 2.1, subset the matrix so that only correlations within
    # the given range are used (correlations can be both positive and negative)
    
    countries <- colnames(m2)
    
    n <- network(m2)
    la <- network.layout.fruchtermanreingold(n, NULL)
    
    country_points <- data.frame(x=la[,1],y=la[,2], country=countries);
      
    event.data <- event_data("plotly_click")
      edge_shapes <- list()
      n <- length(country_points$country)

      for(i in 1:n) {
        for(j in 1:n) {
          
          if (!is.na(m2[i,j]) && m2[i,j] > 0) {
            if (is.null(event.data)){
              width <- 1
            
              # Problem 2.2, highlight the selected link by increasing its
              # width to some larger than 1 value
              edge_shape = list(type = "line",
              line = list(color = "#030303", width = width))
              edge_shape[['x0']] = country_points$x[i]
              edge_shape[['y0']] = country_points$y[i]
              edge_shape[['x1']] = country_points$x[j]
              edge_shape[['y1']] = country_points$y[j]
              edge_shapes <- c(edge_shapes, list(edge_shape))
            }
            else{
              if (attr(m2,"dimnames")[[1]][i]==event.data$x && attr(m2,"dimnames")[[2]][j]==event.data$y){
              
                width <- 3
                
                # Problem 2.2, highlight the selected link by increasing its
                # width to some larger than 1 value
                edge_shape = list(type = "line",
                line = list(color = "#030303", width = width))
                edge_shape[['x0']] = country_points$x[i]
                edge_shape[['y0']] = country_points$y[i]
                edge_shape[['x1']] = country_points$x[j]
                edge_shape[['y1']] = country_points$y[j]
                edge_shapes <- c(edge_shapes, list(edge_shape))
              }
              
              else {
                width <- 1
                
                # Problem 2.2, highlight the selected link by increasing its
                # width to some larger than 1 value
                edge_shape = list(type = "line",
                line = list(color = "#030303", width = width))
                edge_shape[['x0']] = country_points$x[i]
                edge_shape[['y0']] = country_points$y[i]
                edge_shape[['x1']] = country_points$x[j]
                edge_shape[['y1']] = country_points$y[j]
                edge_shapes <- c(edge_shapes, list(edge_shape))
              }
            }
          }
        }
      }
      
      ax <- list(
        showgrid = FALSE,
        showticklabels = FALSE,
        showline = FALSE,
        zeroline = FALSE
      );
      
      net <- plot_ly(country_points, x=~x, y=~y, text=~country, xaxis=ax, yaxis=ax)

      p <- layout(net,
                  shapes=edge_shapes, xaxis=ax, yaxis=ax)
    });

  output$plot <- renderPlotly({
    # renderPlotly() also understands ggplot2 objects!
    # Problem 2.1, subset the correlation matrix and display it as a heatmap
    output$plot <- renderPlotly({
      tc <- t[ , !(names(t) %in% c("week"))]
      m <- abs(cor(tc))
      m2 <- m
      m2[abs(m)<input$slide] <- 0
      plot_ly(x=rownames(m2), y=colnames(m2), z = m2, type = "heatmap")
    
    })
    
    output$click <- renderPrint({
      d <- event_data("plotly_click")

      if (is.null(d) == T) return (NULL);
      
      #updateSelectInput(session, "country", selected=country_points$country[d$pointNumber + 1]);
      if (is.null(d)) "Click events appear here (double-click to clear)" else d
    })
  })
};

shinyApp(ui, server, options=list(height=1000))
```


